\section{Introduction}\label{sec:introduction}

Frequency estimation sits on the critical path of relaying, under-frequency
load shedding, disturbance monitoring, and converter supervision in
low-inertia grids. In inverter-based resource (IBR) systems, the problem is
not limited to slow electromechanical drift. Harmonics, abrupt phase jumps,
out-of-band components, and short ringdown segments can arrive on the same
timescale as the protection decision itself~\cite{Milano2018LowInertia,
Pinheiro2025,Ponce2025}. Recent IBR-focused studies also show that converter
control interactions and mixed disturbances can reshape the observable
frequency response itself~\cite{Bernal2025,Zoghby2024}. An estimator that
performs well in a textbook sinusoidal test can still generate large transient
error or non-negligible trip-risk when these effects overlap.

The estimator families used in practice already reveal a tension between
accuracy, latency, and disturbance tolerance. Windowed spectral methods reject
steady harmonic content but pay observation delay~\cite{Belega2014,
PlatasGarza2011}. PLL variants are light enough for embedded deployment but can
slip under abrupt phase discontinuities~\cite{Golestan2017,Wright2019}.
Kalman-filter variants can track fast transients when the state model remains
valid~\cite{Ferrero2020EKF,Chamorro2021}. Reduced-order and data-driven methods
offer another point on the cost-accuracy curve, but their behavior under
relay-oriented composite stress is even less standardized~\cite{Netto2018KoopmanKF,
Ting2022}. What remains unclear is how these families reorder once the
benchmark moves beyond isolated standard tests.

This paper treats that question as the main result. Rather than forcing a
single winner across all conditions, we ask three practical questions: which
estimators lead the standard disturbance families, which IBR stress cases
reorder the ranking, and which methods remain attractive once CPU cost and
trip-risk are considered alongside RMSE.

The benchmark analyzed here covers 16 estimators and a broad scenario set
that mixes IEEE- and NERC-style disturbances with IBR-specific harmonics,
ringdown, and multi-event stress. The main finding is as follows. The dominant
estimator changes with the disturbance family: the second-order
generalized-integrator PLL (SOGI-PLL) leads ramps, the extended Kalman filter
(EKF) leads magnitude steps and out-of-band interference, the robust adaptive
EKF (RA-EKF) leads modulation, and the IBR multi-event case splits RMSE and
trip-risk leadership between the phase-locked loop (PLL) and interpolated DFT
(IpDFT). That scenario dependence, rather than any single ranking, is the
main message of the paper.
