\begin{abstract}
Frequency estimators used in relay and monitoring functions face a harder
problem in inverter-based resource (IBR) grids than in classical synchronous
systems: phase jumps, harmonics, ringdown, and out-of-band components can
appear in the same event. Standard compliance-style benchmarks still emphasize
isolated disturbances, so they say little about how estimator rankings move
under composite stress. This paper analyzes a Monte Carlo benchmark covering 16
estimators, 34 benchmark scenarios, and 60 runs per estimator--scenario
pair. It spans
IEEE-style ramps, magnitude steps, modulation, phase jumps, out-of-band
interference, and IBR-specific harmonics, ringdown, and multi-event stress.
We report RMSE, peak error, trip-risk duration, settling time, and CPU time
per sample. The results do not support a single best estimator. The
second-order generalized-integrator PLL (SOGI-PLL) gives the lowest mean RMSE
on ramp families (19.4\,mHz), the extended Kalman filter (EKF) leads
magnitude-step and out-of-band cases (0.388\,mHz and 13.6\,mHz), and the robust
adaptive EKF (RA-EKF) leads modulation (17.8\,mHz). In the IBR multi-event
case, the phase-locked loop (PLL) attains the lowest RMSE (173.5\,mHz) while
the interpolated DFT (IpDFT) yields the lowest trip-risk there ($0.0439$\,s). Nine
different estimators win at least one scenario on RMSE, and in 9 of the 12
benchmark scenarios with a unique trip-risk winner, that winner differs from the RMSE
winner. The hardest cases are the IBR multi-event sequence (826\,mHz median
RMSE across estimators), the $60^\circ$ phase jumps, and noisy ringdown
scenarios. The error range spans nearly three orders of magnitude, showing that
isolated tests miss scenario dependence that matters
for protection design.
\end{abstract}
