\subsection{Related Work}
\label{sec:related}

Standards and many published benchmark studies still center on isolated
disturbance classes. IEEE C37.118.1 and IEC/IEEE 60255-118-1 define dynamic
tests around ramps, magnitude steps, modulation, and a limited set of phase
events~\cite{IEEE_C37_118,IEC_IEEE_60255_118_1_2018}. Ting et al. extend that
logic to low-inertia frequency measurement methods and show clear trade-offs
between tracking speed and disturbance rejection even under a carefully chosen
set of simulated events~\cite{Ting2022}. Wright et al. further show that phase
steps can produce false ROCOF signatures, which is precisely the kind of
measurement ambiguity that simple compliance views do not resolve~\cite{Wright2019}.

Recent IBR-focused studies make the gap sharper rather than smaller.
Low-inertia surveys and system studies show that converter penetration changes
both the spatial distribution and the time scale of frequency dynamics~\cite{Milano2018LowInertia,
Pinheiro2025,Ponce2025}. Related benchmark proposals for hybrid islanded or
grid-forming systems emphasize interaction-driven behavior, but they do not
provide a broad estimator comparison over standard PMU tests, composite IBR
stress, trip-risk, and compute cost in one frame~\cite{Bernal2025,Zoghby2024}.
That is the limitation this paper addresses. A method that performs well on a
clean standard test may fail under conditions that the test never probes.

The estimator families themselves are not the missing piece. Window-based
methods trade harmonic rejection against observation delay~\cite{Belega2014,
PlatasGarza2011}. PLL variants are cheap enough for embedded deployment but can
lose phase alignment under abrupt discontinuities~\cite{Golestan2017,Wright2019}.
Kalman-filter variants can improve transient tracking when their state model
matches the event~\cite{Ferrero2020EKF,Chamorro2021}. Reduced-order and
data-driven methods add further cost-accuracy trade-offs~\cite{Netto2018KoopmanKF}.
The missing piece addressed here is the comparison protocol: a broad benchmark
that keeps
scenario family, trip-risk, and compute cost in view at the same time. The
contribution of this paper is therefore not a new estimator but a benchmark
study that preserves standard families, adds composite IBR stress, and compares
accuracy, transient risk, and compute cost within one reporting framework.
